Analysis plan

### 2023-03-23 02. Design/MSG-P2312-Analysis Plan-1-clusters.rmd
Update MSG-P2312-Analysis Plan-1-clusters.rmd
Added a table defining terms, clarified how teams are defined, started cleaning use of terms, added comments from Jake, added variables to capture colocation, and added new rules-based categorization based on position category

@@

The analysis will be done for staff and teams separately and aggregated
independently. Employees will be categorized. Also, employees with be
grouped into teams based on supervisor. These teams will also be
categorized. The table below defines how we plan to aggregate for staff
and teams.

@@

ADD TEAM CLAFFICATION RULES FOR STAFF AND TEAM ANALYSIS

We will approach grouping staff and teams in two ways. First,
rules-based groups will be formed. For these, definitions of different
types of work will be defined. Second, groups will be formed using
k-means clustering. This unsupervised machine learning method produces a
defined number of groups (or clusters) minimizing the distance between
the members.

@@

For staff-level analysis, the team will be defined by the employee's
supervisor. Changes in supervisors will come from staff moving teams or
an existing team changing supervisors. The size of the team will be
caluculated with the mode of staff with the supervisor over the
observations window. For employees with multiple supervisors, the size
of the team will be the average of the modes of team sizes for each
supervisor.

@@

-   Proportion of days of staff concurrent remote status

-   Proportion of days of staff concurrent onsite status

@@

@@

For the rules based approach, the portion of working days remote will be
the key variable to determine what category a staff or team will be in.
Staff will be grouped as follows:

1.  **All remote** - A policy definition for staff that are 100% remote.
2.  **Teleworker** - A policy definition for staff that require two days
    per pay period onsite.
3.  **Primarily remote** - Staff who work remotely above the 95th
    percentile <!-- could be 75th -->
4.  **Mostly remote** - Staff who work remotely between the median and
    the 95th percentile <!-- could be 75th -->
5.  **Mostly onsite** - Staff who work remotely between 5th percentile
    and the median <!-- could be 25th -->
6.  **Primarly onsite** - Staff who work remotely below the 5th
    percentile <!-- could be 75th -->

@@

1.  **Primarily remote** - Teams that work remotely above the 95th
    percentile <!-- could be 75th -->
2.  **Mostly remote** - Teams that work remotely between the median and
    the 95th percentile <!-- could be 75th -->
3.  **Mostly onsite** - Teams that work remotely between 5th percentile
    and the median <!-- could be 25th -->
4.  **Primarily onsite** - Teams that work remotely below the 5th
    percentile <!-- could be 75th -->

@@

For staff clusters, the following variables will be used to define the
clusters:

@@

@@

-   Averages size of team over observation window

@@

@@

    -   Proportion of concurrent remote days for team members

    -   Proportion of concurrent onsite days for team members

>>>>

#### Table `r table_n`: Definitions of teams used in this document

```{r}
tbl_definitions <- tibble("Term" = c(
  "Staff",
  "Employee",
  "Team",
  "Team member",
  "Supervisor",
"Category",
"Rule-based group",
"Cluster"

)
)
```

```{r}
tbl <- 
  tbl_definitions %>% 
  add_column(
    "Description" = c(
  "GSA staff, used when writing more generally",
"GSA staff when writing about employee-level analysis",
"A group of employees when writing about team-level analysis",
"Employee when writing about employee-level analysis",
"GSA staff who supervise a team, used to define who is in a team",
"General team for classifications of employees or teams",
"Category based on some defined rule (i.e. remote worker)",
"Category based on results for cluster analysis"
)
)
```

```{r}
tbl %>% 
  flextable() %>% 
  # add_header_row(values = c("","","Type of analysis", "Type of analysis")) %>% 
  # set_header_labels(Group = "Level of analysis") %>% 
  merge_h(part = "header") %>% 
  merge_v(j = 1) %>% 
  theme_vanilla() %>%
  bold(j = 1) %>% 
  autofit()
```

@@

The analysis will be done for employees and teams separately and aggregated
independently. Employees will be categorized. Also, team members will be defined as belonging to a team based on supervisor. These teams will also be
categorized. The table below defines how we plan to aggregate for staff
and teams.

@@

# CONTINUE REVIEWING TERMS FROM HERE.

# Update based on notes from conversation with Jake 3-23-23

<!-- scaling variables -->
<!--  - use rank -->
<!--  - demean and divide by std deviation -->
<!-- compare clustering results -->
<!-- check variable distribution -->
<!-- check pairwise plots of variables -->

We will categorize employees and teams in two ways. First,
rules-based groups will be formed. For these, definitions of different
types of work will be defined based on the proportion of work done remotely, by policy rules, and by position category. Second, groups will be formed using
k-means clustering. This unsupervised machine learning method produces a
defined number of groups (or clusters) minimizing the distance between
the members.

@@

-   Proportion of working days co-located with at least one other team
    member

-   Proportion of working days co-located with supervisor

-   Proportion of working days co-located with entire team

-   Proportion of working days entire team is remote

For staff-level analysis, the team will be defined by the employee's
supervisor. Changes in supervisors will come from staff moving teams or
an existing team changing supervisors. The size of the team will be
calculated with the mode of staff with the supervisor over the
observations window. For employees with multiple supervisors, the size
of the team will be the average of the modes of team sizes for each
supervisor.

@@

@@

-   Proportion of working days at least two team member co-located

-   Proportion of working days supervisor is co-located with at least
    one team member

-   Proportion of working days entire team is co-located

-   Proportion of working days entire team is remote

@@

For the rules-based approach, policy definitions, the position category
(as defined by `position_category_description`), and the portion of
working days remote will be determine what category a staff or team will
be in. Staff will be grouped as follows:

1.  **All remote** - A policy definition for staff that are 100% remote
2.  **Teleworker** - A policy definition for staff that require two days
    per pay period onsite.
3.  **Onsite Flexible** - Defined by `position_category_description` in
    the roster data
4.  **Offsite** - Defined by `position_category_description` in the
    roster data
5.  **Onsite Required** - Defined by `position_category_description` in
    the roster data
6.  **Primarily remote** - Staff who work remotely above the 95th
    percentile <!-- could be 75th -->
7.  **Mostly remote** - Staff who work remotely between the median and
    the 95th percentile <!-- could be 75th -->
8.  **Mostly onsite** - Staff who work remotely between 5th percentile
    and the median <!-- could be 25th -->
9.  **Primarily onsite** - Staff who work remotely below the 5th
    percentile <!-- could be 75th -->

@@

1.  **All remote** - All team members are100% remote
2.  **Teleworker** - All team members work two days per pay period
    onsite.
3.  **Onsite Flexible** - All team members are *Onsite Flexible* as
    defined by `position_category_description` in the roster data
4.  **Offsite** - All team members are *Offsite* as defined by
    `position_category_description` in the roster data
5.  **Onsite Required** - All team members are *Onsite Required* as
    defined by `position_category_description` in the roster data
6.  Mixed position categories - Team members are different position
    categories as as defined by `position_category_description` in the
    roster data
7.  **Primarily remote** - Teams that work remotely above the 95th
    percentile <!-- could be 75th -->
8.  **Mostly remote** - Teams that work remotely between the median and
    the 95th percentile <!-- could be 75th -->
9.  **Mostly onsite** - Teams that work remotely between 5th percentile
    and the median <!-- could be 25th -->
10. **Primarily onsite** - Teams that work remotely below the 5th
    percentile <!-- could be 75th -->

@@

For employee clusters, the following variables will be used to define
the clusters:

@@

-   Modal size of team over observation window

-   Proportion of working days co-located with at least one other team
    member

-   Proportion of working days co-located with supervisor

-   Proportion of working days co-located with entire team

-   Proportion of working days entire team is remote

@@

-   Modal size of team over observation window

@@

    -   Proportion of working days co-located with at least one other
        team member

    -   Proportion of working days co-located with supervisor

    -   Proportion of working days co-located with entire team

    -   Proportion of working days entire team is remote

@@
